---
title: "VIS2128 ZHANG Portfolio"
author: "Lamei Zhang"
date: "10/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Loading libraries

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = 'hide'}
library(leaflet)
library(sf)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggspatial)
library(htmlwidgets)
library(tidytransit)
```


##Maps of Boston & Cambridge

Using various techniques in R and ArcGIS, I display geospatial analysis in Boston and neighboring Cambridge.

###Interactive Map (Assignment #6)

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = 'hide'}
Middlesex_White <- get_decennial(geography = "block",
                          state = "MA", county = "Middlesex",
                          year = 2010,
                          output = "wide",
                          variables = c(tot_pop = 'P008001',
                                        white_pop = 'P005003'),
                          geometry = TRUE)

Middlesex_Hispanic <- get_decennial(geography = "block",
                                    state = "MA", county = "Middlesex",
                                    year = 2010,
                                    output = "wide",
                                    variables = c(tot_pop = 'P008001',
                                                  his_pop = 'P009002'),
                                    geometry = TRUE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = 'hide'}
Middlesex_White <- Middlesex_White %>%
  mutate(pct_white = white_pop / tot_pop) %>%
  filter(tot_pop > 0)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = 'hide'}
MA_state_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000.0001016002 +y_0=750000 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}

MA_White_palette <- colorNumeric(c("green", "red"), Middlesex_White$pct_white)

White_map3 <- leaflet(Middlesex_White) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(color = ~MA_White_palette(pct_white), stroke = FALSE, fillOpacity = 0.7,
              highlightOptions = highlightOptions(fillColor = "darkorange", fillOpacity = 0.9),
              label = Middlesex_White$NAME,
              popup = paste("Total population: ", Middlesex_White$tot_pop, "<br/>",
                            "White (non-Hispanic) population: ", Middlesex_White$bl_pop, " (", 
                            round(Middlesex_White$pct_white * 100, 1), "%)", sep = "")) %>%
    addLegend("bottomright", pal = MA_White_palette, values = ~pct_white,
    title = "Percent of population<br/>identifying as White alone",
    labFormat = labelFormat(suffix = "%",
                            transform = function(x) 100 * x),
    opacity = 1)

White_map3

```

###Walkability of Philadelphia (Assignment #4)




###Raster Display of Philadelphia (Assignment #5)




###Georeferencing historic Chelsea (Assignment #2)




###Calculating Density in Cambridge (Assignment #3)

